\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-05-12 paper.qmd
Initial images for digit1 analysis

>>>>

<!-- PaCMAP layout with author's suggested parameter choice-->
```{r}
#| warning: false
#| echo: false

## Import data
training_data_mnist <- read_rds("data/mnist/mnist_10_pcs_of_digit_1.rds")
training_data_mnist <- training_data_mnist |>
  mutate(ID = 1:NROW(training_data_mnist))

pacmap_minst <- read_rds("data/mnist/mnist_pacmap.rds")
pacmap_minst_scaled <- gen_scaled_data(data = pacmap_minst)$scaled_nldr

position_df <- pacmap_minst_scaled |> 
  filter(ID %in% c(6475, 3311, 6489, 2347, 387, 
                          5239, 728, 517, 6369, 3055, 6345, 1761)) |>
  mutate(position = c(5, 9, 7, 8, 4, 11, 2, 6, 12, 10, 1, 3))

pacmap_plot_mnist <- pacmap_minst_scaled |> 
  ggplot(aes(x = PaCMAP1, 
             y = PaCMAP2)) + 
  geom_point(alpha=0.1) + 
  geom_text(data = position_df, 
            aes(x = PaCMAP1, y = PaCMAP2, label = position), 
            colour = "red", 
            size = 5)  
```

<!--Render the images in different selected location in 2D layout-->
```{r}
#| warning: false
#| echo: false

## Data with pixel values
mnist_data <- read_rds("data/mnist/mnist_digit_1.rds")

pixels_gathered <-  mnist_data |>
  mutate(instance = row_number()) |>
  gather(pixel, value, -Label, -instance) |>
  tidyr::extract(pixel, "pixel", "(\\d+)", convert = TRUE) |>
  mutate(pixel = pixel - 2, x = pixel %% 28, y = 28 - pixel %/% 28) |>
  filter(instance %in% c(3710, 2391, 2385, 5030, 6475,
                         7679, 3568, 1312, 3311, 1097,
                         3552, 7853, 6489, 7689, 6690,
                        1380, 6057, 2347, 5946, 3355,
                        4175, 3997, 5378, 387, 1854,
                        614, 3079, 1762, 5239, 3723,
                        5748, 728, 7419, 7794, 6233,
                        33, 2485, 5998, 318, 1761,
                        5690, 165, 517, 6935, 2682,
                        4962, 2264, 5563, 6369, 559,
                        5188, 4849, 2666, 3448, 3055,
                        3120, 6869, 6345, 4470, 7147)) |>
  mutate(instance.labs =  case_when(
    instance %in% c(3710, 2391, 2385, 5030, 6475) ~ "1",
    instance %in% c(7679, 3568, 1312, 3311, 1097) ~ "2",
    instance %in% c(3552, 7853, 6489, 7689, 6690) ~ "3",
    instance %in% c(1380, 6057, 2347, 5946, 3355) ~ "4",
    instance %in% c(4175, 3997, 5378, 387, 1854) ~ "5",
    instance %in% c(614, 3079, 1762, 5239, 372) ~ "6",
    instance %in% c(5748, 728, 7419, 7794, 6233) ~ "7",
    instance %in% c(5690, 165, 517, 6935, 2682) ~ "8",
    instance %in% c(4962, 2264, 5563, 6369, 559) ~ "9",
    instance %in% c(5188, 4849, 2666, 3448, 3055) ~ "10",
    instance %in% c(3120, 6869, 6345, 4470, 7147) ~ "11",
    .default = "12"
    ))

## To generate labels
instance.labs <- as.character(rep(1:12, each = 5))

img_sample <- pixels_gathered |> 
  ggplot(aes(x, y, fill = value)) +  
  geom_tile() +
  facet_wrap(~ factor(instance, levels = c(3710, 2391, 2385, 5030, 6475,
                         7679, 3568, 1312, 3311, 1097,
                         3552, 7853, 6489, 7689, 6690,
                        1380, 6057, 2347, 5946, 3355,
                        4175, 3997, 5378, 387, 1854,
                        614, 3079, 1762, 5239, 3723,
                        5748, 728, 7419, 7794, 6233,
                        33, 2485, 5998, 318, 1761,
                        5690, 165, 517, 6935, 2682,
                        4962, 2264, 5563, 6369, 559,
                        5188, 4849, 2666, 3448, 3055,
                        3120, 6869, 6345, 4470, 7147)), ncol = 5) +
  scale_fill_continuous_sequential(palette = "Grays") +
  theme(strip.background = element_blank(),
        strip.text.x = element_blank())
```

<!--PaCMAP param: n_components=2, n_neighbors=10, init=random, MN_ratio=0.9, FP_ratio=2.0-->
```{r}
#| echo: false
#| fig-cap: "(a) 2D layout from PaCMAP applied for the digit 1 of the MNIST dataset. Is this a best representation of the original data?, and (b) Images of the handwritten digit 1. The angle of the digit 1 varies along non-linear structure."
#| label: fig-pacmapauthor
#| fig-pos: H

pacmap_plot_mnist + img_sample +
  plot_annotation(tag_levels = 'a')
```

### 2024-05-12 paper.tex
Initial images for digit1 analysis

>>>>

\begin{figure}[H]

\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pacmapauthor-1.pdf}

\caption{\label{fig-pacmapauthor}(a) 2D layout from PaCMAP applied for
the digit 1 of the MNIST dataset. Is this a best representation of the
original data?, and (b) Images of the handwritten digit 1. The angle of
the digit 1 varies along non-linear structure.}
